---
title: "data 605 wk 10 assignment"
author: "Jun Pan"
date: "April 4, 2019"
output:
  html_document: default
  pdf_document: default
---

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. 
 
Find the probability that he wins 8 dollars before losing all of his money if 
 
 
(a) he bets 1 dollar each time (timid strategy). 


```{r}
trials <- 1000
out_of_jail <- 0
for (i in 1:trials) {
  bank <- 1
  while (bank>0 & bank<8) {
    bank <- bank + sample(c(-1,1),1, replace=TRUE, prob=c(0.6,0.4))
  }
  out_of_jail<-out_of_jail+(bank>=8)
}
paste0("Smith got out of jail ",out_of_jail,
      " times in ",trials,
      " simulations with probability of success ",out_of_jail/trials,".")
```
So the chance for smith got of jail using strategy a is 0.027.




(b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy). 


```{r}
P <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0,
              q, 0, p, 0, 0, 0, 0, 0, 0,
              q, 0, 0, 0, p, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              q, 0, 0, 0, 0, 0, 0, 0, p,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 0,
              0, 0, 0, 0, 0, 0, 0, 0, 1), nrow = 9, ncol = 9, byrow = TRUE)
rownames(P) <- 0:8
colnames(P) <- 0:8
P

Q <- P[c(2:8), c(2:8)]
I <- diag(nrow = nrow(Q))
Q

N <- solve(I - Q)
N

R <- P[c(2:8), c(1, 9)]
R

# compute B

B <- N %*% R
B
round(B[1,2], 4)
```

So using B strategy, the change to get out of jial is 0.064


(c) Which strategy gives Smith the better chance of getting out of jail? 

B strategy gives Simth the better chance of getting out of jail.
